Calculating the Solvent Exposed Surface Area of Proteins with Intersecting Spheres

### 2020-06-06 Paper.tex
More Edits

The Shrake-Rupley Algorithm (Shrake and Rupley, 1973) is the standard way of calculating SASA by approximating the surface as a discrete set of voxels instead of a continuous surface. This can lead to inaccuracies, but the gains in speed are usually worth it. Prior exact algorithms could also fail if the protein surface was not continuous (i.e. an internal cavity is present) or the atomic radii are too small. This algorithm, once optimized by Legrand and Merz (1993) became a standard SASA calculator. Comparing this algorithm with the analytic SASA calculator MSEED (Perrot et al.) in calculating the SASA of the protein BPTI showed Shrake-Rupley to be roughly six times faster than MSEED and not subject to the surface continuity problems that can trouble the analytic calculator. The method works by representing the surface of sphere $a_i$ by an array of $l$ points evenly distributed on the surface of a sphere of radius $r_i$. The molecular surface area can be then written as integral of didio $\Psi\Phi\Xi\alpha^2$

@@

%The authors make use of a power diagram for their model, which is another name for the structure of spheres shown in Figure 2. Just like we mentioned previously, the power diagram consists of power cells. [Repetitive]

@@

\section{Application of ALPHAMOL to Molecular Modeling}
To test the time cost of \textsc{alphamol} (the computational implement of the method described in $\S3$) we took 250 PDB structures from the RCSB database (listed in the supplemental) and calculated the number of polar atoms buried in the structure that are not making hydrogen bonds. This calculation indirectly uses a SASA calculator to determine if atoms close to the surface are truly buried. If they have any SASA that is nonzero, the atom is assumed to be making a hydrogen bond with the implicit solvent. We compared the time taken to complete the buried unsatisfied polars calculation for the \textsc{alphamol} and Shrake-Rupley SASA calculators. We found that \textsc{alphamol} is slower, but it is more accurate.

@@

\bibitem{Levitt} 
Bryant, R., Edelsbrunner, H., Koehl, P., Levitt, M. The Area Derivative of a Space-Filling Diagram. Discrete \& Computational Geometry. 32 (2004), 293-308.

>>>>

The Shrake-Rupley Algorithm (Shrake and Rupley, 1973) is the standard way of calculating SASA by approximating the surface as a discrete set of voxels instead of a continuous surface. This can lead to inaccuracies, but the gains in speed are usually worth it. Prior exact algorithms could also fail if the protein surface was not continuous (i.e. an internal cavity is present) or the atomic radii are too small. This algorithm, once optimized by Legrand and Merz (1993) became a standard SASA calculator. Comparing this algorithm with the analytic SASA calculator MSEED (Perrot et al.) in calculating the SASA of the protein BPTI showed Shrake-Rupley to be roughly six times faster than MSEED and not subject to the surface continuity problems that can trouble the analytic calculator. The method works by representing the surface of sphere $a_i$ by an array of $l$ points evenly distributed on the surface of a sphere of radius $r_i$. The molecular surface area can be then written as integral of 

@@

@@

\section{Application of ALPHAMOL to Molecular Modeling}
To test the time cost of \textsc{alphamol} (the computational implement of the method described in $\S3$) we took 100 PDB structures from the RCSB database (listed in the supplemental) and calculated the number of polar atoms buried in the structure that are not making hydrogen bonds. This calculation indirectly uses a SASA calculator to determine if atoms close to the surface are truly buried. If they have any SASA that is nonzero, the atom is assumed to be making a hydrogen bond with the implicit solvent. We compared the time taken to complete the buried unsatisfied polars calculation for the \textsc{alphamol} and Shrake-Rupley SASA calculators. We found that \textsc{alphamol} is slower, but it is more accurate.

@@

\bibitem{PDB} 
H.M. Berman, J. Westbrook, Z. Feng, G. Gilliland, T.N. Bhat, H. Weissig, I.N. Shindyalov, P.E. Bourne.
(2000) The Protein Data Bank Nucleic Acids Research, 28: 235-242.
